The first prediction of Black hole mass function and its evolution by ET

### 2019-05-17 GW_BH_function/manuscript_GW_constrain_BHmass/main.tex
Update by Kai

\shorttitle{Black hole mass function and redshift evolution by ET}
\shortauthors{Marian et al.}

@@

%\author%[0000-0002-4359-5994]
%{Kai Liao}
%\affiliation{School of Science, Wuhan University of Technology, Wuhan 430070, China}

@@

\begin{abstract}
%discover of GW
%Measure of BH mass and thus fit the BH mass.
%The discovery of double compact objects by GW provide a the measurement of their BH masses. Recent works by LIGO group have constrained 10 GW event and constrain the BHs. The ET would detect more GW events with DL and BH mass up to redshift $z\sim10$, and thus enable one to constrain the mass function and study how it evolution with redshift. In this work, test the ability of 1000 GWs events by ET. We tested, 1. how well are they able to recover the BH mass function as power law distribution. 2. If the GW events are able to tell a redshifted-related BH mass function.
We demonstrate the ability of future gravitational-wave (GW) measured by third generation detector Einstein Telescope (ET) to infer the slope of the BH mass function (BHMF) and its evolution with redshift. %The measurement of BH binary systems enable the inference of the BH mass (\mone\ and \mtwo\ of each component) together with the luminosity distance (\dl).
We performed the Monte Carlo approach to simulate the measurements of BH binaries GW signals as detected by ET, including the BH masses and their luminosity distance (\dl). Following the previous works, we consider the mass of  primary black hole in each binary and model the BHMF as a function of the power law with slope parameter as $\alpha$. We take into account the bias that would be raised by the uncertainty of the measurement and the selection effect and find that a thousand GW events measure by ET ($\sim1\%$ amount of the yearly detection rate) could measure the $\alpha$ to $3\%$ accuracy level. Furthermore, we investigate a scenario, by the first time, that  $\alpha$ is evolving with redshift as $\alpha(z) = \alpha_0 + \alpha_1\frac{z}{1+z}$. Our result is that, taking the \dl\ as redshift estimator, with a thousand GW events one is capable of obtaining an unbiased $\alpha_1$ at the uncertainty level of $\pm0.5$. The ability of known the evolution scenario of BHMF would help to understand the origin of the BHs and how BH binaries formed at the different stage of the Universe.

@@

\section{Introduction} \label{sec_intro}
As first identified as a solution to Einstein's field equations by Schwarzschild in 1916 \citep{Schwarzschild1999}, black holes (BHs) have been predicted to remain a bunches of fundamental questions in unifying GR with quantum physics \citep{Hawking1976, Giddings2017}. The mass of BHs is known to cover a wide range from stellar-mass to supermassive ($\sim10^{10} M_{\odot}$). The recent discovery of gravitational-wave (GW) measurement is a substantial evidence of stellar-mass BHs \citep{Abbott2016}, while supermassive BHs are thought to existe in the centers of almost all the galaxies \citep{Lynden-Bell1969, Kormendy1995}. More recently, the first image of supermassive BH in the center of the giant elliptical galaxy M87 has been reconstructed by radio-wave observations \citep{Alberdi2019}. Nevertheless, in theory, it is still unclear of how the BHs are formed \citep{Fryer1999, Fryer2001, Mirabel2016}. The prediction for the number and mass distribution of stellar-mass BHs in the Universe is still needed to be more clear.

The recent GW detection activities have brought us to a new era of GW astronomy \citep[e.g.,][]{Abbott2016, Abbott2016_sum, Abbott2018}. To date, the LIGO and Virgo have achieved two successful runs (O1 and O2) and now is performing the third run (O3). In the first run O1 (from September 12th, 2015 to January 19th, 2016), the first catalog of binary BH (BBH) inspiral systems have been measured including three BBH systems. In the O2 run (from November 30th, 2016 to August 25th, 2017), a total of seven binary black hole mergers have been detected\footnote{Together with the first detection of GW from detection of binary neutron star inspiral system.}. In the next decade, the volume of the events is expected to be increasing rapidly with the improvements in the detector sensitivities.

The GWs provide a direct way to hear the inspiralling BBH systems, enable one to derive their information including mass, spin, luminosity distance\citep{Abbott2017phy}. This brings us the opportunity to measure the population of BHs \citep{Abbott2018b} and helps to answer some fundamental questions such as the cosmography\citep{Liao2017, Ding2019, Cai2017}, the GW speed\citep{Fan2017, Collett2017}, the strong lensing effects~\citep{Ola2013, Biesiada2014, Ding2015}. 
The previous work \citet{Kovetz2017PhRvD} has investigated that further LIGO measurements from thousands of BBHs can be taken to infer the power law index $\alpha$ at 10\% accuracy. More recently, the LIGO collaboration has used ten BBH merger events and constrained the BHMF as power law index to $\alpha~=~1.6\substack{+1.5\\-1.7}$ (90\% credibility) \citep{Abbott2018b}. In this study, we pay extra attention on simulating the GW events that could be measured by third-generation gravitational waves detector, Einstein Telescope (ET) \citep{Abernathy2011} based Monte Carlo (MC) approach. This provides us a mock BBHs catalog with the realistic dataset to exam their ability to constrain the BHMFs, taking into the effects by the data noise level, selection bias in a more realistic way. Moreover, the ET would detect the GW event at distant Universe up to $z\sim17$ \citep{Abernathy2011}. The wide redshift range of the events enable us to exam whether it is possible to detect the $\alpha$ as a function of redshift.

@@

To randomly generate the BH masses, we follow the previous works \citep{Kovetz2017PhRvD, Abbott2018b, Fishbach2018} and assume the \mone\ follows a power law distribution with hard cut at both maximum and minimum mass:
 \begin{equation} \label{equ_powlaw}
p(m_1|\alpha, M_{max}, M_{min}) = m_1^{\alpha} \mathcal{H}(m_1-M_{min}) \mathcal{H}(M_{max}-m_1),
 \end{equation}
where $\mathcal{H}$ is the Heaviside step function. Then, the secondary mass, \mtwo, is fix as uniform between $[M_{min}, m_1]$. Note that, we only take the \mone\ to reconstruct the BHMF, thus the assumption of the distribution for \mtwo\ actually does not affect the inference for the shape of BHMF.
The generation of these parameters would determine the value of $\Theta$ and ${\cal M}_0$ in Equation~\ref{SNR}. We combine them with their redshift to realize the $\rho$ of each BBH system. We collect the events which have $\rho > \rho_0 = 8$, meaning that those events with $\rho < 8$ are to faint to be detected.

@@

\subsection{Selection Effect}\label{sec_likelihood_sf}
The observation the has tendency to discover more significant events, known as Malmquist bias. The GW systems with higher values \mone\ trend to stronger signals and thus have a higher possibility to be detected. If not well taking into account, the final BHMFs would be biased the high mass end in particular.

@@

\vspace{1cm}
\section{Result}\label{sec_result}
%We present the result in this section.
%One thousand of data points.
%An appendix to introduce the correction for the skewness for the $\eta$.
We fit the mock data to the BHMF model to infer the distribution of the best-fitted parameter. To infer a non-biased distribution, we adopt the realization approach. The detail of the simulation if the following. In each realization, we simulate a thousand GW events and infer the best-fit parameters using minimization. We do not perform the realizations until the inferred best-fit parameters have converged and informative.

@@

\vspace{1cm}
\section{Conclusion \& Discussion} \label{sec_summary}
The third-generation gravitational waves detector, Einstein Telescope, is very powerful and capable of detecting $\sim10^5$ GW events per year, with redshift up to $z\sim17$. In this study, we investigated how the detections of the merger of BBHs improve our knowledge of the black hole mass function.

>>>>

\shorttitle{Black hole mass function and redshift evolution by ET}
\shortauthors{Ding et al.}

@@

\author%[0000-0002-4359-5994]
{Kai Liao}
\affiliation{School of Science, Wuhan University of Technology, Wuhan 430070, China}

@@

\begin{abstract}
%discover of GW
%Measure of BH mass and thus fit the BH mass.
%The discovery of double compact objects by GW provide a the measurement of their BH masses. Recent works by LIGO group have constrained 10 GW event and constrain the BHs. The ET would detect more GW events with DL and BH mass up to redshift $z\sim10$, and thus enable one to constrain the mass function and study how it evolution with redshift. In this work, test the ability of 1000 GWs events by ET. We tested, 1. how well are they able to recover the BH mass function as power law distribution. 2. If the GW events are able to tell a redshifted-related BH mass function.
We demonstrate the ability of future gravitational-waves (GW) measured by third generation detector Einstein Telescope (ET) to infer the slope of the BH mass function (BHMF) and its evolution with redshift. %The measurement of BH binary systems enable the inference of the BH mass (\mone\ and \mtwo\ of each component) together with the luminosity distance (\dl).
We performed the Monte Carlo approach to simulate the measurements of binary BH (BBH) inspiral systems signals as detected by ET, including the BH masses and their luminosity distances (\dl). Following the previous works, we consider the mass of  primary black hole in each binary and model the BHMF as a function of the power law with slope parameter as $\alpha$. We take into account the bias that would be raised by the uncertainty of the measurement and the selection effect and find that a thousand GW events measured by ET ($\sim1\%$ amount of the yearly detection rate) could measure the $\alpha$ to $3\%$ accuracy level. Furthermore, we investigate a scenario, that  $\alpha$ is evolving with redshift as $\alpha(z) = \alpha_0 + \alpha_1\frac{z}{1+z}$. Our result is that, taking the \dl\ as redshift estimator, with a thousand GW events one is capable of obtaining an unbiased $\alpha_1$ at the uncertainty level of $\pm0.5$. The ability of knowing the evolution scenario of BHMF would help to understand the origin of the BHs and how BH binaries formed at different stage of the Universe.

@@

\section{Introduction} \label{sec_intro}
As first identified as a solution to Einstein's field equations by Schwarzschild in 1916 \citep{Schwarzschild1999}, black holes (BHs) have been predicted to remain a bunches of fundamental questions in unifying GR with quantum physics \citep{Hawking1976, Giddings2017}. The mass of BHs is known to cover a wide range from stellar-mass to supermassive ($\sim10^{10} M_{\odot}$). The recent discovery of gravitational-wave (GW) measurement is a substantial evidence of stellar-mass BHs \citep{Abbott2016}, while supermassive BHs are thought to exist in the centers of almost all the galaxies \citep{Lynden-Bell1969, Kormendy1995}. More recently, the first image of supermassive BH in the center of the giant elliptical galaxy M87 has been reconstructed by radio-wave observations \citep{Alberdi2019}. Nevertheless, in theory, it is still unclear of how the BHs are formed \citep{Fryer1999, Fryer2001, Mirabel2016}. The number and mass distribution of stellar-mass BHs in the Universe is still needed to be more clear.

The recent GW detection activities have brought us to a new era of astronomy \citep[e.g.,][]{Abbott2016, Abbott2016_sum, Abbott2018}. To date, the LIGO and Virgo have achieved two successful runs (O1 and O2) and now is performing the third run (O3). In the first run O1 (from September 12th, 2015 to January 19th, 2016), the first catalog of binary BH (BBH) inspiral systems have been measured including three BBH systems. In the O2 run (from November 30th, 2016 to August 25th, 2017), a total of seven binary black hole mergers have been detected\footnote{Together with the first detection of GW from detection of binary neutron star inspiral system.}. In the next decade, the volume of the events is expected to be increasing rapidly with the improvements in the detector sensitivities.

The GWs provide a direct way to hear the inspiralling BBH systems, enable one to derive their information including mass, spin, luminosity distance\citep{Abbott2017phy}. This brings us the opportunity to not only measure the population of BHs \citep{Abbott2018b}, but also answer some fundamental questions such as the cosmography\citep{Liao2017, Ding2019, Cai2017}, the GW speed\citep{Fan2017, Collett2017}, the strong lensing effects~\citep{Ola2013, Biesiada2014, Ding2015}.

Focusing on the BH mass function (BHMF), the recent work \citet{Kovetz2017PhRvD} has investigated that further LIGO measurements from thousands of BBHs can be taken to infer the  BHMF as power law and constrain the slope parameter $\alpha$ at 10\% accuracy. More recently, the LIGO collaboration has used ten BBH merger events and constrained the BHMF as power law index to $\alpha~=~1.6\substack{+1.5\\-1.7}$ (90\% credibility) \citep{Abbott2018b}. In this study, we pay extra attention on simulating the GW events that could be measured by third-generation gravitational wave detector, Einstein Telescope (ET) \citep{Abernathy2011} based Monte Carlo (MC) approach. This provides us a mock BBHs catalog with the realistic dataset to exam their ability to constrain the BHMFs, taking into the effects by the data noise level, selection bias in a more realistic way. Moreover, the ET would detect the GW event at distant Universe up to $z\sim17$ \citep{Abernathy2011}. The wide redshift range of the events enable us to exam whether it is possible to detect the $\alpha$ as a function of redshift.

@@

To randomly generate the BH masses, we follow the previous works \citep{Kovetz2017PhRvD, Abbott2018b, Fishbach2018} and assume the \mone\ follows a power law distribution with hard cut at both maximum and minimum mass:
 \begin{equation} \label{equ_powlaw}
p(m_1|\alpha, M_{max}, M_{min}) = m_1^{\alpha} \mathcal{H}(m_1-M_{min}) \mathcal{H}(M_{max}-m_1),
 \end{equation}
where $\mathcal{H}$ is the Heaviside step function. Then, the secondary mass, \mtwo, is fixed as uniform between $[M_{min}, m_1]$. Note that, we only take the \mone\ to reconstruct the BHMF, thus the assumption of the distribution for \mtwo\ actually does not affect the inference for the shape of BHMF.
The generation of these parameters would determine the value of $\Theta$ and ${\cal M}_0$ in Equation~\ref{SNR}. We combine them with their redshift to realize the $\rho$ of each BBH system. We collect the events which have $\rho > \rho_0 = 8$, meaning that those events with $\rho < 8$ are to faint to be detected.

@@

\subsection{Selection Effect}\label{sec_likelihood_sf}
The observation has tendency to discover more significant events, known as Malmquist bias. The GW systems with higher values \mone\ trend to stronger signals and thus have a higher possibility to be detected. If not well taking into account, the final BHMFs would be biased the high mass end in particular.

@@

\vspace{1cm}
\section{Result}\label{sec_result}
%We present the result in this section.
%One thousand of data points.
%An appendix to introduce the correction for the skewness for the $\eta$.
We fit the mock data to the BHMF model to infer the distribution of the best-fitted parameter. To infer a non-biased distribution, we adopt the realization approach. The detail of the simulation is the following. In each realization, we simulate a thousand GW events and infer the best-fit parameters using minimization. We do not perform the realizations until the inferred best-fit parameters have converged and informative.

@@

\vspace{1cm}
\section{Conclusion \& Discussion} \label{sec_summary}
The third-generation gravitational wave detector, Einstein Telescope, is very powerful and capable of detecting $\sim10^5$ GW events per year, with redshift up to $z\sim17$. In this study, we investigated how the detections of the merger of BBHs improve our knowledge of the black hole mass function.
